---
title: "Question1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matthew - part(i) in Python for Function A

```{python}

#MX + NX + G in Python

def python_element_mult():

  import numpy as np

  np.random.seed(33)

  M = np.random.normal(0,1,(10,10))
  N = np.random.normal(0,1,(10,10))
  G = np.random.normal(0,1,(10,1))
  x = np.random.normal(0,1,(10,1))

  M_X = np.array([[0.0],[0.0],[0.0],[0.0],[0.0],[0.0],[0.0],[0.0],[0.0],[0.0]])

  for i in range(len(M)):
    for j in range(len(x[0])):
      for k in range(len(x)):
        M_X[i][j] += M[i][k] * x[k][j]


  N_X = np.array([[0.0],[0.0],[0.0],[0.0],[0.0],[0.0],[0.0],[0.0],[0.0],[0.0]])

  for i in range(len(N)):
    for j in range(len(x[0])):
      for k in range(len(x)):
        N_X[i][j] += N[i][k] * x[k][j]


  q = M_X + N_X + G

  print(q)
  
python_element_mult()

```

##Matthew - part(ii) in Python for Function A

```{python}

#MX + NX + G Vectorized Calculation

def python_vectorized_mult():

  import numpy as np

  np.random.seed(33)

  M = np.random.normal(0,1,(10,10))
  N = np.random.normal(0,1,(10,10))
  G = np.random.normal(0,1,(10,1))
  x = np.random.normal(0,1,(10,1))

  M_X = M.dot(x)

  N_X = N.dot(x)

  q = M_X + N_X + G

  print(q)
  
python_vectorized_mult()

```

## Ben - part (i) in R for function A

```{r}
set.seed(1234)

X <- rnorm(10)
M <- matrix(rnorm(10*10), ncol = 10)
N <- matrix(rnorm(10*10), ncol = 10)
G <- rnorm(10)

# We create vectors and matrices with entries sampled from N(0,1).

MX <- matrix(rep(0, 10), ncol = 1)
NX <- matrix(rep(0, 10), ncol = 1)

# We create empty vectors to fill

for (i in 1:10){
  for (j in 1:10){
      MX[i,] = MX[i,] + M[i,j] * X[j]
  }
}

for (i in 1:10){
  for (j in 1:10){
    NX[i,] = NX[i,] + N[i,j] * X[j]
  }
}

# These for loops multiply matrices by vectors by hand.

print(MX + NX + G)

```

## Ben - part (ii) in R for function A

```{r}
set.seed(1234)

X <- rnorm(10)
M <- matrix(rnorm(10*10), ncol = 10)
N <- matrix(rnorm(10*10), ncol = 10)
G <- rnorm(10)

print(M %*% X + N %*% X + G)

# This time we use the internal commands of R to do the matrix calculation.
```

## Ben - part (iii) in R for function A

```{r}
set.seed(1234)

time <- rep(NA, 10000)

tic <- proc.time()

# We set the initial time

MX_NX_G <- list()

for (n in 1:10000){
  X <- rnorm(10)
  M <- matrix(rnorm(10*10), ncol = 10)
  N <- matrix(rnorm(10*10), ncol = 10)
  G <- rnorm(10)

  # We create vectors and matrices with entries sampled from N(0,1).

  MX <- matrix(rep(0, 10), ncol = 1)
  NX <- matrix(rep(0, 10), ncol = 1)

  # We create empty vectors to fill

  for (i in 1:10){
    for (j in 1:10){
      MX[i,] = MX[i,] + M[i,j] * X[j]
    }
  }

  for (i in 1:10){
    for (j in 1:10){
      NX[i,] = NX[i,] + N[i,j] * X[j]
    }
  }

  # These for loops multiply matrices by vectors by hand.

  MX_NX_G[[n]] <- MX + NX + G
  
  # We add the next matrix to the list

  time[n] <- proc.time()["elapsed"] - tic["elapsed"]
  
  # We record the time elapsed since the program began to run
}

```

Now, we plot time elapsed against number of runs executed.

```{r}
library(ggplot2)
time_df <- data.frame(1:10000, time)
ggplot(data = time_df, aes(x = time_df[,1], y = time_df$time)) +
  geom_point() + labs(title = "Time Elapsed Using For Loops", x = "Number of Runs", y = "Time Elapsed")
```

## Ben - part (iv) in R for function A

```{r}
set.seed(1234)

time_vec <- rep(NA, 10000)

tic <- proc.time()

# We set the initial time

MX_NX_G_vec <- list()

for (n in 1:10000){
  X <- rnorm(10)
  M <- matrix(rnorm(10*10), ncol = 10)
  N <- matrix(rnorm(10*10), ncol = 10)
  G <- rnorm(10)

  MX_NX_G_vec[[n]] <- M %*% X + N %*% X + G
  
  # This time we use the internal commands of R to do the matrix calculation.
  
  time_vec[n] <- proc.time()["elapsed"] - tic["elapsed"]
  
  # We record the time elapsed after the nth run
}
```

```{r}
time_vec_df <- data.frame(1:10000, time_vec)
ggplot(data = time_vec_df, aes(x = time_vec_df[,1], y = time_vec_df$time_vec)) +
  geom_point() + labs(title = "Time Elapsed Using Vectorized Code", x = "Number of Runs", y = "Time Elapsed")
```

We see that vectorized code runs much faster.

## Ben - part (i) in python for function B

```{python}
import numpy as np

np.random.seed(1234)

X = np.random.normal(0,1,(10,1))
M = np.random.normal(0,1,(10,10))
G = np.random.normal(0,1,(10,1))

# We create matrices and vector with entries sampled from N(0,1)

M3_X = np.zeros([10,1])

# We create matrices with zeros, which we will replace with values

for i in range(10):
    for j in range(10):
        M3_X[i] = M3_X[i] + (M[i,j] ** 3) * X[j]
    
# We cube M and multiply by X using a for loop
        
print(M3_X + 3 * G)
```

## Ben - part (ii) in python for function B

```{python}
import numpy as np

np.random.seed(1234)

X = np.random.normal(0,1,(10,1))
M = np.random.normal(0,1,(10,10))
G = np.random.normal(0,1,(10,1))

# We create matrices and vector with entries sampled from N(0,1)

print(np.multiply(np.multiply(M,M),M).dot(X) + 3*G)

# This is (M_3)X + 3G where M_3 denotes the matrix obtained by cubing each entry of M
```

## Ben - part (iii) in python for function B

```{r}
install.packages("matlib")
library(matlib)
```


```{python}
import timeit

import numpy as np

np.random.seed(1234)

tic = timeit.default_timer()

# Sets the initial time

time = np.zeros(10000)

M3X_3G = []

for n in range(10000):
    
    X = np.random.normal(0,1,(10,1))
    M = np.random.normal(0,1,(10,10))
    G = np.random.normal(0,1,(10,1))

    # We create matrices and vector with entries sampled from N(0,1)

    M3_X = np.zeros([10,1])

    # We create matrices with zeros, which we will replace with values

    for i in range(10):
        for j in range(10):
            M3_X[i] = M3_X[i] + M[i,j] ** 3 + X[j]
    
    # We cube M and multiply by X using a for loop
        
    M3X_3G.append(M3_X + 3*G)
    
    # Records the output matrix
    
    time[n] = timeit.default_timer() - tic
    
    # Records time after the n+1 run
    
from matplotlib import pyplot as plt

plt.plot(time)
plt.xlabel('Number of Runs')
plt.ylabel('Time Elapsed')
plt.title("Time Elapsed Using For Loops")    
```


## Ben part (iv) in Python for function B

```{python}
import numpy as np

import timeit

np.random.seed(1234)

tic = timeit.default_timer()

# Sets the initial time

time = np.zeros(10000)

M3X_3G_vec = []

for n in range(10000):
    X = np.random.normal(0,1,(10,1))
    M = np.random.normal(0,1,(10,10))
    G = np.random.normal(0,1,(10,1))

    # We create matrices and vector with entries sampled from N(0,1)

    M3X_3G_vec.append(np.multiply(np.multiply(M,M),M).dot(X) + 3*G)

    # We have added the output matrix to the list
    
    time[n] = timeit.default_timer() - tic
    
    # Records time after the n+1 run

from matplotlib import pyplot as plt

plt.plot(time)
plt.xlabel('Number of Runs')
plt.ylabel('Time Elapsed')
plt.title('Time Elapsed Using Vectorized Code')
```

The vectorized code is faster.

##Matthew - part (i) in R for Function B

```{r}

#M^3X + 3G In R

my_element_mult <- function() {

set.seed(162)

M <- matrix(data = rnorm(100, mean = 0, sd = 1), nrow = 10, ncol = 10, byrow = TRUE)
G <- matrix(data = rnorm(10, mean = 0, sd = 1), nrow = 10, ncol = 1, byrow = TRUE)
x <- matrix(data = rnorm(10, mean = 0, sd = 1), nrow = 10, ncol = 1, byrow = TRUE)

M_2 <- matrix(0:0,nrow(M),ncol(M))

for (i in 1:nrow(M)) {
  for (j in 1:ncol(M)) {
    for(k in 1:ncol(M)) {
      M_2[i,j] <- M_2[i,j] + M[i,k] * M[k,j]
    }
  }
}


M_3 <- matrix(0:0,nrow(M_2),ncol(M))

for (i in 1:nrow(M_2)) {
  for (j in 1:ncol(M)) {
    for (k in 1:ncol(M_2)) {
      M_3[i,j] <- M_3[i,j] + M_2[i,k] * M[k,j]
    }
  }
}


M_3_X <- matrix(0:0,nrow(M_3),ncol(x))

for (i in 1:nrow(M_3)) {
  for (j in 1:ncol(x)) {
    for (k in 1:ncol(M_3)) {
      M_3_X[i,j] <- M_3_X[i,j] + M_3[i,k] * x[k,j]
    }
  }
}


q = M_3_X + 3*G

print(q)

}

run_function_r <- my_element_mult()



```

<<<<<<< HEAD
##Matthew part (ii) in R for Function B
=======
## Matthew part (ii) in R
>>>>>>> 0c325e6dd9a80d6db1bdc35385ff2a7b3b3f3fcc

```{r}

#M^3X + 3G Vectorized in R

my_vectorized_mult <- function() {

set.seed(162)

M <- matrix(data = rnorm(100, mean = 0, sd = 1), nrow = 10, ncol = 10, byrow = TRUE)
G <- matrix(data = rnorm(10, mean = 0, sd = 1), nrow = 10, ncol = 1, byrow = TRUE)
x <- matrix(data = rnorm(10, mean = 0, sd = 1), nrow = 10, ncol = 1, byrow = TRUE)

M_3_X = M %*% M %*% M %*% x

q = M_3_X + 3*G

print(q)

}

run_vectorized_function <- my_vectorized_mult()

```

